A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950

### 2022-02-12 getdata4.Rmd
add socStatus and urban

@@

```{r readpers}
per <- pers[(pers$frsbobtyp == 2) & (pers$df == 0), ]
per$frsbobtyp <- per$df <- per$f2id <- per$m2id <- NULL
per$frsbobdatmin <- per$frsbobdatmax <- per$utdatmax <- per$utdatmin <- NULL
per <- per[(per$foddat > 18950000), ]
dn <- per$doddat < 1
per$doddat[dn] <- per$foddat[dn] + 40000
## Fix foddat:
yr <- per$foddat %% 10000 == 0
per$foddat[yr] <- per$foddat[yr] + 101 # As early as possible
mon <- per$foddat %% 100 == 0
per$foddat[mon] <- per$foddat[mon] + 1 # Ditto
## Fix doddat:
yr <- per$doddat %% 10000 == 0
per$doddat[yr] <- per$doddat[yr] + 1231 # As late as possible
mon <- per$doddat %% 100 == 0
per$doddat[mon] <- per$doddat[mon] + 28 # Ditto (guard for February ...)
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
nabir <- is.na(per$birthdate)
per$foddat[nabir] <- per$foddat[nabir] - 2
per$doddat[nabir] <- per$doddat[nabir] - 2
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate))
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
## Fix doddat:
per$doddat[is.na(per$deathdate)] <- per$doddat[is.na(per$deathdate)] - 1
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
##
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
oj <- per$exit < 0
per$doddat[oj] <- per$doddat[oj] + 10000
per$deathdate <-  as.Date(as.character(per$doddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
per$event <- per$exit <= 365
per$exit[per$exit == 0] <- 0.25
per$exit[per$exit > 365] <- 365
```

@@

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "none"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "low"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "enter", "exit", "event")]
saveRDS(per, file = "data/births4.rds")
```

@@

bume$region <- "ume"
bskel$region <- "ske"
bsten$region <- "sten"
infdat <- rbind(bume, bskel, bsten)
infdat$birthmonth <- ceiling((infdat$birthdate - floor(infdat$birthdate)) * 12)
infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
infdat <- infdat[order(infdat$id, infdat$enter), c("subreg", "id", "sex", "enter", "exit", "event", 
                                                   "birthmonth", "period", "socBranch", "socStatus", "illeg",
                                                   "parity", "hisclass", tempvars)]
## New addition: (28 Jan 2022)
##infdat <- rc(infdat) # rc from 'skum'
##n <- NROW(infdat)

>>>>

Added *urban* on 12 February 2022.

@@

```{r readpers}
per <- pers[(pers$frsbobtyp == 2) & (pers$df == 0), ]
per$urban <- per$fboort %in% c(274, 1208)
per$frsbobtyp <- per$df <- per$f2id <- per$m2id <- NULL
per$frsbobdatmin <- per$frsbobdatmax <- per$utdatmax <- per$utdatmin <- NULL
per <- per[(per$foddat > 18950000), ]
dn <- per$doddat < 1
per$doddat[dn] <- per$foddat[dn] + 40000
## Fix foddat:
yr <- per$foddat %% 10000 == 0
per$foddat[yr] <- per$foddat[yr] + 101 # As early as possible
mon <- per$foddat %% 100 == 0
per$foddat[mon] <- per$foddat[mon] + 1 # Ditto
## Fix doddat:
yr <- per$doddat %% 10000 == 0
per$doddat[yr] <- per$doddat[yr] + 1231 # As late as possible
mon <- per$doddat %% 100 == 0
per$doddat[mon] <- per$doddat[mon] + 28 # Ditto (guard for February ...)
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
nabir <- is.na(per$birthdate)
per$foddat[nabir] <- per$foddat[nabir] - 2
per$doddat[nabir] <- per$doddat[nabir] - 2
##
per$birthdate <- as.Date(as.character(per$foddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate))
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
## Fix doddat:
per$doddat[is.na(per$deathdate)] <- per$doddat[is.na(per$deathdate)] - 1
per$deathdate <- as.Date(as.character(per$doddat), format = "%Y%m%d")
##
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
oj <- per$exit < 0
per$doddat[oj] <- per$doddat[oj] + 10000
per$deathdate <-  as.Date(as.character(per$doddat), format = "%Y%m%d")
per$exit <- as.numeric(difftime(per$deathdate, per$birthdate, units = "days"))
per$event <- per$exit <= 365
per$exit[per$exit == 0] <- 0.25
per$exit[per$exit > 365] <- 365
```

@@

```{r fromobs}
bobs <- obs[obs$starttyp == 2 & obs$id %in% per$id, ]
indx <- match(per$id, bobs$id)
per$socStatus <- bobs$socStatus[indx]
per$socBranch <- bobs$socBranch[indx]
per$hisclass <- bobs$hisclass[indx]
per$nofrs <- bobs$nofrs[indx]
per <- per[!is.na(per$nofrs), ]
per$subreg <- factor(per$nofrs)
levels(per$subreg) <- c(rep("ume", 2), 
                        rep("bjur", 4), 
                        rep("sten", 3))
##per <- per[(per$subreg != "sten"), ] # NOTE: Now Stensele station is included
per$subreg <- factor(per$subreg)
per$enter <- 0
per$sex <- factor(per$kon, labels = c("boy", "girl"))
per$birthdate <- toTime(per$birthdate)
per$illeg <- as.numeric(per$ab == 2)
per$parity <- per$paritet_g
per$paritet_g <- per$paritet_k <- per$doddat <- per$frsbosdat <- NULL
per$fid <- per$fboort <- per$ab <- per$fyrkfd <- per$fyrkind <- NULL
per$region <- per$kon <- per$fodhfrs <- per$frsbobdat <- per$frsbostyp <- NULL
per$fbonmn <- per$fboind <- per$fyrkrel <- per$fyrketxt <- NULL
per$foddat <- as.Date(as.character(per$foddat), format = "%Y%m%d")
## NEW:
per$socBranch <- as.character(per$socBranch)
per$socBranch[is.na(per$socBranch)] <- "none"
per$socBranch <- factor(per$socBranch)
per$socStatus <- as.character(per$socStatus)
per$socStatus[is.na(per$socStatus)] <- "low"
per$socStatus <- factor(per$socStatus)
per <- per[, c("id", "sex", "birthdate", "mid", "foddat", "parity", "illeg", 
               "subreg", "socBranch", "socStatus", "hisclass", "urban", "enter", "exit", "event")]
saveRDS(per, file = "data/births4.rds")
```

@@

bume$region <- "ume"
bskel$region <- "ske"
bsten$region <- "sten"
infdat <- rbind(bume, bskel, bsten)
infdat$birthmonth <- ceiling((infdat$birthdate - floor(infdat$birthdate)) * 12)
infdat$period <- cut(infdat$birthdate, c(1895, 1914, 1935, 1951), dig.lab = 5)
infdat <- infdat[order(infdat$id, infdat$enter), c("subreg", "id", "sex", "enter", "exit", "event", 
                                                   "birthmonth", "period", "socBranch", "socStatus", "illeg",
                                                   "parity", "hisclass", "urban", tempvars)]
## New addition: (28 Jan 2022)
##infdat <- rc(infdat) # rc from 'skum'
##n <- NROW(infdat)

### 2022-02-12 paper4.Rmd
add socStatus and urban

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
#- \usepackage[tablesonly]{endfloat}
#- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature, but effects are different in neonatal and postneonatal settings."
---

@@

@@

    inftab <- toTpch(Surv(enter, exit, event) ~ subreg + sex + period + socBranch + socStatus + parity + 
                         extemp + extemp.1 + emeantemp + summer + ageIvl, data = all,
                     cuts = c(0, 3, 7, 14, 28, 56, 84, 145, 207, 286, 365) / 365)
    tabneo <- inftab[inftab$ageIvl == "neo", ]
    tabpost <- inftab[inftab$ageIvl == "postneo", ]
    tabpost$age <- factor(tabpost$age)
    tabneo$age <- factor(tabneo$age)
    saveRDS(tabneo, file = "data/tabneo.rds")
    saveRDS(tabpost, file = "data/tabpost.rds")
    saveRDS(inftab, file = "data/inftab.rds")
}else{
    inftab <- readRDS("data/inftab.rds") #The rest are read later...
}
```

```{r joint, cache = TRUE}
fit.all <- tpchreg(oe(event, exposure) ~  strata(ageIvl) * 
                       (extemp + extemp.1 + emeantemp + socBranch + period), 
                   data = inftab, time = age)
dr.all <- drop1(fit.all, test = "Chisq")
##system.time(tabneo.fit <- tpchreg(oe(event, exposure) ~ extemp + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabneo, time = age))
##system.time(tabpost.fit <- tpchreg(oe(event, exposure) ~ extemp + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabpost, time = age))
```

@@

```{r formneo}
formneo <- as.formula("oe(event, exposure) ~  extemp + extemp.1 + emeantemp + socBranch + period +
                     subreg + sex + parity")
tabneo <- readRDS("data/tabneo.rds")
```

@@

```{r neoanalw, results = "asis", cache = FALSE}
ltx2(fit.neow, dr = dr, caption = "Neonatal mortality, October to March.", 
     label = "tab:neoanalw", keep = 3)
```

@@

```{r neoanals, results = 'asis', cache = FALSE}
ltx2(fit.neos, dr = dr.neos, caption = "Neonatal mortality, April to September.", 
     label = "tab:neoanals", keep = 3)
```

@@

```{r setformw}
form <- as.formula("oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + subreg + sex + parity")
form.sb <-  as.formula("oe(event, exposure) ~ (extemp + extemp.1 + emeantemp) * socBranch + subreg + sex + parity")
form.sr <- as.formula("oe(event, exposure) ~ (extemp + extemp.1 + emeantemp) * subreg + socBranch + sex + parity") 
tabpost <- readRDS("data/tabpost.rds")
```

@@

```{r postanalwp, results='asis', cache = TRUE}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + extemp + emeantemp + sex + socBranch + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

@@

#### Sex and temperature

```{r postanalwpsex, results='asis', cache = TRUE}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + extemp + emeantemp + sex + socBranch + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

fit.postwsex <- tpchreg(oe(event, exposure) ~ sex * (extemp + extemp.1 + emeantemp), 
                     data = tabpost[!tabpost$summer, ], time = age)
dr.postwsex <- drop1(fit.postwsex, test = "Chisq")
```

@@

```{r postanalwsex, cache = FALSE}
round(dr.postwsex, 4)
#tbl(round(dr.postw, 4), caption = "Interaction tests with temperature and period, postneonatal ##mortality.", rownames = TRUE)
```

Obviously no interactions between temperature and sex.

@@

```{r postw1srpu, cache = TRUE}
fit.postw1sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1895,1914]" & tabpost$subreg == "ume", ], time = age)
dr.postw1sru <- drop1(fit.postw1sru, test = "Chisq")
```

```{r postw1sru, results = "asis"}
ltx2(fit.postw1sru, dr = dr.postw1sru, caption = "Postneonatal mortality, Umeå, October to March 1895-1913.",
     label = "tab:postw1sru", keep = 3)
```

```{r postw1srps, cache = TRUE}
fit.postw1srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1895,1914]" & tabpost$subreg != "ume", ], time = age)
dr.postw1srs <- drop1(fit.postw1srs, test = "Chisq")
```

```{r postw1srs, results = "asis"}
ltx2(fit.postw1srs, dr = dr.postw1srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1895-1913.", 
     label = "tab:postw1srs", keep = 3)
```

@@

```{r postw2, results = "asis", cache = FALSE}
ltx2(fit.postw2, dr = dr.postw2, caption = "Postneonatal mortality, October to March 1914-1934. Adjusted for sex, parity, and subregion.", label = "tab:postw2", keep = 3)
```

@@

```{r postw2srpu, cache = TRUE}
fit.postw2sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1914,1935]" & tabpost$subreg == "ume", ], time = age)
dr.postw2sru <- drop1(fit.postw2sru, test = "Chisq")
```

```{r postw2sru, results = "asis"}
ltx2(fit.postw2sru, dr = dr.postw2sru, caption = "Postneonatal mortality, Umeå, October to March 1914-1934.",
     label = "tab:postw2sru", keep = 3)
```

```{r postw2srps, cache = TRUE}
fit.postw2srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1914,1935]" & tabpost$subreg != "ume", ], time = age)
dr.postw2srs <- drop1(fit.postw2srs, test = "Chisq")
```

```{r postw2srs, results = "asis"}
ltx2(fit.postw2srs, dr = dr.postw2srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1914-1934.", 
     label = "tab:postw2srs", keep = 3)
```

@@

```{r postw3p, results = "asis"}
fit.postw3 <- tpchreg(oe(event, exposure) ~ extemp.1 + emeantemp + socBranch + subreg + sex + parity, 
                      data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]", ], time = age)
##fit.postw3 <- tpchreg(form, data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]", ], time = age)
dr.postw3 <- drop1(fit.postw3, test = "Chisq")
```

```{r postw3, results = "asis", cache = FALSE}
ltx2(fit.postw3, dr = dr.postw3, caption = "Postneonatal mortality, October to March 1935-1950. Adjusted for sex, parity, and subregion.", label = "tab:postw3", keep = 3)
```

@@

```{r postw3srpu, cache = TRUE}
fit.postw3sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]" & tabpost$subreg == "ume", ], time = age)
dr.postw3sru <- drop1(fit.postw3sru, test = "Chisq")
```

```{r postw3sru, results = "asis"}
ltx2(fit.postw3sru, dr = dr.postw3sru, caption = "Postneonatal mortality, Umeå, October to March 1935-1950.",
     label = "tab:postw3sru", keep = 3)
```

```{r postw3srps, cache = TRUE}
fit.postw3srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]" & tabpost$subreg != "ume", ], time = age)
dr.postw3srs <- drop1(fit.postw3srs, test = "Chisq")
```

```{r postw3srs, results = "asis"}
ltx2(fit.postw3srs, dr = dr.postw3srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1895-1913.", 
     label = "tab:postw3srs", keep = 3)
```

@@

fit.posts <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + period + 
                        subreg + parity + sex, 
                    data = tabpost[tabpost$summer, ], time = age)
dr <- drop1(fit.posts, test = "Chisq")
```

```{r postanals, results='asis', cache = FALSE}
ltx2(fit.posts, dr = dr, caption = "Postneonatal mortality, April to September. Adjusted for sex, parity, time period, and subregion.", label = "tab:postanals", keep = 3)
```

>>>>

---
title: "A biometric analysis of infant mortality and temperature, northern Sweden 1895-1950"
author: "Göran Broström and Tommy Bengtsson"
date: "`r Sys.time()`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  bookdown::word_document2:
    toc: no
    toc_depth: 2
  bookdown::html_document2:
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
citation_package: natbib
classoption: titlepage
bibliography: bio.bib
titlepage: yes
biblio-style: apalike
documentclass: article
fontsize: 11pt
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{graphicx}
- \usepackage{a4wide}
- \usepackage[tablesonly]{endfloat}
- \renewcommand{\efloatseparator}{\mbox{}}
abstract: "The effect of extreme temperatures on infant mortality in the Umeå and Skellefteå regions 1895-1950 is studied in a biometric analysis setting. More precisely, the effect of climate and weather, measured by temperature, average and extremes, on infant mortality is investigated. It turns out that climate (average) is more important than weather (extremes), low average temperatures are more important than temporary dips in temperature, but effects are different in neonatal and postneonatal settings."
---

@@

**socStatus** Social status *within* the given working branch.

@@

    inftab <- toTpch(Surv(enter, exit, event) ~ subreg + sex + period + socBranch + socStatus + urban + parity + 
                         extemp + extemp.1 + emeantemp + summer + ageIvl, data = all,
                     cuts = c(0, 3, 7, 14, 28, 56, 84, 145, 207, 286, 365) / 365)
    tabneo <- inftab[inftab$ageIvl == "neo", ]
    tabpost <- inftab[inftab$ageIvl == "postneo", ]
    tabpost$age <- factor(tabpost$age)
    tabneo$age <- factor(tabneo$age)
    saveRDS(tabneo, file = "data/tabneo.rds")
    saveRDS(tabpost, file = "data/tabpost.rds")
    saveRDS(inftab, file = "data/inftab.rds")
}else{
    inftab <- readRDS("data/inftab.rds") #The rest are read later...
}
```

```{r joint, cache = TRUE}
fit.all <- tpchreg(oe(event, exposure) ~  strata(ageIvl) * 
                       (extemp + extemp.1 + emeantemp + socBranch + socStatus + period + urban), 
                   data = inftab, time = age)
dr.all <- drop1(fit.all, test = "Chisq")
##system.time(tabneo.fit <- tpchreg(oe(event, exposure) ~ extemp + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabneo, time = age))
##system.time(tabpost.fit <- tpchreg(oe(event, exposure) ~ extemp + emeantemp + highTemp + lowTemp + socBranch + socStatus, 
  ##               data = tabpost, time = age))
```

@@

```{r formneo}
formneo <- as.formula("oe(event, exposure) ~  extemp + extemp.1 + emeantemp + socBranch + socStatus + period + urban +
                     subreg + sex + parity")
tabneo <- readRDS("data/tabneo.rds")
```

@@

```{r neoanalw, results = "asis", cache = FALSE}
ltx2(fit.neow, dr = dr, caption = "Neonatal mortality, October to March.", 
     label = "tab:neoanalw", keep = NULL)
```

@@

```{r neoanals, results = 'asis', cache = FALSE}
ltx2(fit.neos, dr = dr.neos, caption = "Neonatal mortality, April to September.", 
     label = "tab:neoanals", keep = NULL)
```

@@

```{r setformw}
form <- as.formula("oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + subreg + sex + parity")
form.sb <-  as.formula("oe(event, exposure) ~ (extemp + extemp.1 + emeantemp) * socBranch + socStatus + urban + subreg + sex + parity")
form.sr <- as.formula("oe(event, exposure) ~ (extemp + extemp.1 + emeantemp) * subreg + socBranch + socStatus + urban + sex + parity") 
tabpost <- readRDS("data/tabpost.rds")
```

@@

```{r postanalwp, results='asis', cache = TRUE}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + extemp + emeantemp + sex + socBranch + socStatus + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

@@

#### Sex, urban,  and temperature

```{r postanalwpsex, results='asis', cache = TRUE}
##tpneo <- toTpch(Surv(enter, exit, event) ~ lowTemp + highTemp + extemp + emeantemp + sex + socBranch + socStatus + illeg + parity
  ##              + subreg + period, data = infdat, cuts = c(0, 1))
##winter <- gtpneo[gtpneo$summer == 0, ]
##winter$season <- factor(winter$season)

fit.postwsex <- tpchreg(oe(event, exposure) ~ (sex + urban) * (extemp + extemp.1 + emeantemp), 
                     data = tabpost[!tabpost$summer, ], time = age)
dr.postwsex <- drop1(fit.postwsex, test = "Chisq")
```

@@

```{r postanalwsex, cache = FALSE}
round(dr.postwsex, 4)
#tbl(round(dr.postw, 4), caption = "Interaction tests with temperature and period, urban, postneonatal ##mortality.", rownames = TRUE)
```

Obviously no interactions between temperature and sex, and urban.

@@

```{r postw1srpu, cache = TRUE}
fit.postw1sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1895,1914]" & tabpost$subreg == "ume", ], time = age)
dr.postw1sru <- drop1(fit.postw1sru, test = "Chisq")
```

```{r postw1sru, results = "asis"}
ltx2(fit.postw1sru, dr = dr.postw1sru, caption = "Postneonatal mortality, Umeå, October to March 1895-1913.",
     label = "tab:postw1sru", keep = NULL)
```

```{r postw1srps, cache = TRUE}
fit.postw1srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1895,1914]" & tabpost$subreg != "ume", ], time = age)
dr.postw1srs <- drop1(fit.postw1srs, test = "Chisq")
```

```{r postw1srs, results = "asis"}
ltx2(fit.postw1srs, dr = dr.postw1srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1895-1913.", 
     label = "tab:postw1srs", keep = NULL)
```

@@

```{r postw2, results = "asis", cache = FALSE}
ltx2(fit.postw2, dr = dr.postw2, caption = "Postneonatal mortality, October to March 1914-1934. Adjusted for sex, parity, and subregion.", label = "tab:postw2", keep = NULL)
```

@@

```{r postw2srpu, cache = TRUE}
fit.postw2sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1914,1935]" & tabpost$subreg == "ume", ], time = age)
dr.postw2sru <- drop1(fit.postw2sru, test = "Chisq")
```

```{r postw2sru, results = "asis"}
ltx2(fit.postw2sru, dr = dr.postw2sru, caption = "Postneonatal mortality, Umeå, October to March 1914-1934.",
     label = "tab:postw2sru", keep = NULL)
```

```{r postw2srps, cache = TRUE}
fit.postw2srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1914,1935]" & tabpost$subreg != "ume", ], time = age)
dr.postw2srs <- drop1(fit.postw2srs, test = "Chisq")
```

```{r postw2srs, results = "asis"}
ltx2(fit.postw2srs, dr = dr.postw2srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1914-1934.", 
     label = "tab:postw2srs", keep = NULL)
```

@@

```{r postw3p, results = "asis"}
fit.postw3 <- tpchreg(oe(event, exposure) ~ extemp.1 + emeantemp + socBranch + socStatus + urban + subreg + sex + parity, 
                      data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]", ], time = age)
##fit.postw3 <- tpchreg(form, data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]", ], time = age)
dr.postw3 <- drop1(fit.postw3, test = "Chisq")
```

```{r postw3, results = "asis", cache = FALSE}
ltx2(fit.postw3, dr = dr.postw3, caption = "Postneonatal mortality, October to March 1935-1950. Adjusted for sex, parity, and subregion.", label = "tab:postw3", keep = NULL)
```

@@

```{r postw3srpu, cache = TRUE}
fit.postw3sru <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]" & tabpost$subreg == "ume", ], time = age)
dr.postw3sru <- drop1(fit.postw3sru, test = "Chisq")
```

```{r postw3sru, results = "asis"}
ltx2(fit.postw3sru, dr = dr.postw3sru, caption = "Postneonatal mortality, Umeå, October to March 1935-1950.",
     label = "tab:postw3sru", keep = NULL)
```

```{r postw3srps, cache = TRUE}
fit.postw3srs <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + sex + parity, 
                        data = tabpost[(!tabpost$summer) & tabpost$period == "(1935,1951]" & tabpost$subreg != "ume", ], time = age)
dr.postw3srs <- drop1(fit.postw3srs, test = "Chisq")
```

```{r postw3srs, results = "asis"}
ltx2(fit.postw3srs, dr = dr.postw3srs, caption = "Postneonatal mortality, Skellefteå with inland, October to March 1895-1913.", 
     label = "tab:postw3srs", keep = NULL)
```

@@

fit.posts <- tpchreg(oe(event, exposure) ~ extemp + extemp.1 + emeantemp + socBranch + socStatus + urban + period + 
                        subreg + parity + sex, 
                    data = tabpost[tabpost$summer, ], time = age)
dr <- drop1(fit.posts, test = "Chisq")
```

```{r postanals, results='asis', cache = FALSE}
ltx2(fit.posts, dr = dr, caption = "Postneonatal mortality, April to September. Adjusted for sex, parity, time period, and subregion.", label = "tab:postanals", keep = NULL)
```
